Summary of cell type ontologies in reference files

### 2025-08-18 analyses/cell-type-consensus/exploratory-notebooks/06-update-consensus-scimilarity.Rmd
include extra epithelial cells

# for all cell types we want to look at print out a table of the pairs that give rise to that cell type
celltypes_to_explore |> 
  purrr::map(
    \(celltype){
      lca_df |> 
        dplyr::filter(cl_annotation == {{celltype}}) |> 
        dplyr::select(id1_annotation, id2_annotation, total_lca, cl_annotation) |> 
        unique() |> 
        DT::datatable(rownames = FALSE)
    }
  )
```

@@

```{r}
# now filter to remove the multi LCA matches and add back in the updated assignments for them 
filtered_lca_df <- lca_df |> 
  # first set criteria to keep cell types, > 170, celltypes of interest 
  dplyr::filter(total_descendants <= 170 & !(cl_annotation %in% celltypes_to_exclude) 
                | (cl_annotation == "neuron"
                   # only keep epithelial if not keratinocyte
                | (cl_annotation == "epithelial cell" & id1_annotation != "keratinocyte" & id2_annotation != "keratinocyte"))) |> 
  dplyr::group_by(id1, id2) |> 
  # after filtering, see how many LCAs are left
  dplyr::mutate(new_lca_pair = length(lca)) |> 
   # everything with more than 1 lca gets removed with the exception of HSCs
  dplyr::filter(new_lca_pair <=1 | cl_annotation == "hematopoietic precursor cell") |> 
  dplyr::select(id1, id2, id1_annotation, id2_annotation, lca, cl_annotation, total_descendants) |> 
  # add in pairs with multi LCA and their final match 
  dplyr::bind_rows(multi_lca_keep_df) |> 
  unique()
```

>>>>

# for all cell types we want to look at print out a table of the pairs that give rise to that cell type
celltypes_to_explore |> 
  purrr::map(
    \(celltype){
      lca_df |> 
        dplyr::filter(cl_annotation == {{celltype}}) |> 
        dplyr::select(id1_annotation, id2_annotation, total_lca, cl_annotation) |> 
        unique() |> 
        kable()
    }
  )
```

@@

```{r}
# now filter to remove the multi LCA matches and add back in the updated assignments for them 
filtered_lca_df <- lca_df |> 
  # first set criteria to keep cell types, > 170, celltypes of interest 
  dplyr::filter(total_descendants <= 170 & !(cl_annotation %in% celltypes_to_exclude) 
                | cl_annotation %in% c("neuron","columnar/cuboidal epithelial cell", "endo-epithelial cell")
                   # only keep epithelial if not keratinocyte
                | (cl_annotation == "epithelial cell" & id1_annotation != "keratinocyte" & id2_annotation != "keratinocyte")) |> 
  dplyr::group_by(id1, id2) |> 
  # after filtering, see how many LCAs are left
  dplyr::mutate(new_lca_pair = length(lca)) |> 
   # everything with more than 1 lca gets removed with the exception of HSCs
  dplyr::filter(new_lca_pair <=1 | cl_annotation == "hematopoietic precursor cell") |> 
  dplyr::select(id1, id2, id1_annotation, id2_annotation, lca, cl_annotation, total_descendants) |> 
  # add in pairs with multi LCA and their final match 
  dplyr::bind_rows(multi_lca_keep_df) |> 
  unique()
```
